% Simple template for producing IEEE-format articles using LaTeX.
% Written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% Use at your own risk.  Complaints to /dev/null.

% Make two column with no page numbering.
\documentstyle[twocolumn]{article}
\pagestyle{empty}

%set dimensions of columns and gap between columns
\setlength{\textheight}{8.5in}
\setlength{\columnsep}{2.5pc}
\setlength{\textwidth}{7.0in}
\setlength{\footheight}{0.0in}
\setlength{\topmargin}{0.0in}
\setlength{\oddsidemargin}{-.25in}

%make section titles bold, centered, and a larger font
\makeatletter
\def\section{\@startsection {section}{1}{\z@}{-3.5ex plus -1ex minus 
 -.2ex}{2.3ex plus .2ex}{\Large\bf\centering}}
\makeatother

\begin{document}

%don't want date printed
\date{}

\title{My Wonderful Article in IEEE Format}

%for single author (just remove % characters)
%\author{I. M. Author \\
%  My Department \\
%  My Institute \\
%  My City, STATE, zip}
 
%for two authors (this is what is printed)
\author{\begin{tabular}[t]{c@{\extracolsep{8em}}c}
  I. M. Author	& M. Y. Coauthor \\
 \\
  My Department & Coauthor Department \\
  My Institute & Coauthor Institute \\
  City, STATE~~zipcode	& City, STATE~~zipcode
\end{tabular}}

\maketitle

%I don't know why I have to reset thispagesyle, but otherwise get page numbers
\thispagestyle{empty}

\abstract
This is the abstract of my paper.  It must fit within the size allowed, which
is about 3 inches, including section title.  Extra space can be added as
follows.
\vspace*{1.75in}

\section{Introduction}

Here is my introduction text.  I may want a numbered subsection, which is
done as follows.

\subsection{Previous Work}

Note that size and location of section and subsection titles conforms with
IEEE format.

Throughout I may cite references of the form \cite{key:foo} or
\cite{foo:baz}, and \LaTeX will keep track of numbering.  For those of you
new to the bibliography package, you may have to run the latex process twice
to allow all references to be resolved.

\section{Summary and Conclusions}

This template will get you through the minimum article, i.e. no figures or
equations.  To include those, please refer to your \LaTeX manual.

\subsection*{Acknowledgements}

This is how to do an unnumbered subsection.  Here I thank my colleagues who
know more about \LaTeX than I.

\begin{thebibliography}{9}

\bibitem{key:foo}
I. M. Author, 
``Some Related Article I Wrote,''
{\it Some Fine Journal}, Vol. 17, pp. 1-100, 1987.

\bibitem{foo:baz}
A. N. Expert, 
{\em A Book He Wrote,}
His Publisher, 1989.

\end{thebibliography}

\end{document}
