%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Registration and Hotel Accommodation forms (LaTeX) %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentstyle[11pt]{article}

\topmargin -20mm
\oddsidemargin 0mm
\textheight 240mm
\textwidth 160mm

\parindent 0pt
\parskip 1em

\pagestyle{empty}

\begin{document}

\begin{center}
{\large\bf The Fourth International Workshop on Natural Language Understanding
and Logic Programming}

{\Large\bf Registration Form}
\end{center}

Please complete this form and return it to:

\begin{tabular}{ll}
\hspace*{5mm}	& Yuji Matsumoto (NLULP4)\\
		& Graduate School of Information Science\\
		& Advanced Institute of Science and Technology, Nara\\
		& 8916-5 Takayama-cho, Ikoma-shi, Nara 630-01, Japan\\
		& Phone: +81-7437-2-5240\ Fax:   +81-7437-2-5219\\
		& Email: {\sf matsu@is.aist-nara.ac.jp}
\end{tabular}

{\bf Please type or print in block letters.}

Title: $\Box$ Prof. \ \ $\Box$ Dr. \ \ $\Box$ Ms. \ \ $\Box$ Mr.

Name: \rule[-1mm]{143mm}{.1mm}\\
\hspace*{15mm}(Family Name)\hspace{30mm}(Given
Name)\hspace{30mm}(Other Names)

Institution: \rule[-1mm]{135mm}{.1mm}

Department: \rule[-1mm]{133mm}{.1mm}

Address: \rule[-1mm]{140mm}{.1mm}

\hspace*{15mm} \rule[-1mm]{140mm}{.1mm}

Phone: \rule[-1mm]{66mm}{.1mm}\ \ Fax: \rule[-1mm]{66mm}{.1mm}

Email: \rule[-1mm]{143mm}{.1mm}

{\bf Registration Fee}

\begin{tabular}{lrl}
$\Box$ & 15,000 Yen & (If registration and remmitance completed by August
31, 1993)\\
$\Box$ & 20,000 Yen & (After August 31, 1993)\\
$\Box$ & 8,000 Yen & (Full-time student, certificate requited, no late
fee for students)\\
\end{tabular}

{\bf Remittance}

I have remitted the above sum of total amount in Japanese Yen on
\rule[-1mm]{40mm}{.1mm}(date) through my bank
\rule[-1mm]{40mm}{.1mm}(name of your bank) to the account of NLULP4.

\hspace*{10mm}(Personal check will not be accepted)

Name of Bank: {\bf Nanto Bank, Tomigaoka Branch}\\
Name of Account: {\bf NLULP4}
\hspace{10mm} Account No.: {\bf 82204 (Ordinary Deposit)}

Please send a copy of bank's receipt of the remittance with this form.

\vspace{5mm}
Date: \rule[-1mm]{60mm}{.1mm}\ \  Signature: \rule[-1mm]{60mm}{.1mm}

%%%%%%%%%%%%%
\clearpage

\begin{center}
{\large\bf The Fourth International Workshop on Natural Language Understanding
and Logic Programming}

September 29 -- October 1, 1993\\
New Public Hall, Nara, Japan
\end{center}

\subsection*{Hotel Accommodation}

A number of rooms have been reserved at the following hotels in Nara.
Reservation should be made by completing and returning the attached
{\bf Application Form For Hotel Accommodation}, indicating the name of
the hotel and the number of rooms desired.  No reservation will be
confirmed in the absence of a deposit.  Hotel assignment will be made
on a first-come, first-served basis.

{\bf Deadline of Hotel applications: September 3, 1993}

The daily room charges in Japanese Yen, including tax and service
charges (including breadfast) at specially discounted rates, are as
follows:

\begin{center}
\begin{tabular}{|l|c|c|}\hline
  & \multicolumn{2}{|c|}{Room Charge (in Japanese Yen)}\\ \cline{2-3}
Name of Hotel	  & Single/Bath & Twin/Bath\\ \hline\hline
Nara Hotel	  & 28,000	& 30,600\\ \hline
Nara Garden Hotel & 19,000	& 24,000\\ \hline
Nara Royal Hotel  & 11,845	& 20,600\\ \hline
Wakakusa Hotel	  & 10,300	& 20,600\\ \hline
Hotel Sunroute Nara & 9,800	& 19,000\\ \hline
\end{tabular}
\end{center}

Payment of hotel accommodation can be made either by credit card or by
a bank draft.  \underline{In case of in-land Master Card, please
specify the tie-up card company name.}

In case hotel cancellation, written notification should be sent to JTB
(Japan Travel Bureau).  Deposit will be returned after deducting the
following cancellation charges.

When notification is received by JTB.

\begin{tabular}{ll}
Up to 9 days before the first night of stay: & 1,000 Yen per room\\
8 -- 2 days before & 20\% of the daily room charge\\
One day before or no notice given & 100\% of the daily room charge
\end{tabular}

\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% NLULP4 Hotel Accommodation Form %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\parskip .9em
\begin{center}
{\large The Fourth International Workshop on Natural Language Understanding
and Logic Programming}

{\Large\bf Application Form For Hotel Accommodation}
\end{center}

\vspace{-3mm}
Please complete this form and return it to: (Deadline: September 3, 1993)

\begin{tabular}{ll}
\hspace*{10mm}	& Japan Travel Bureau Kyoto Office\\
		& Convention Department, NLULP\\
		& Higashi shiokoji, Shimogyo, Kyoto 600, Japan\\
		& Fax. +81-75-341-1028\\
		& Phone. +81-75-361-7241
\end{tabular}

{\bf Please type or print in block letters.}

Title: $\Box$ Prof. \ \ $\Box$ Dr. \ \ $\Box$ Mrs. \ \ $\Box$ Ms. \ \ $\Box$ Mr.

Name: \rule[-1mm]{143mm}{.1mm}\\
\hspace*{15mm}(Family Name)\hspace{30mm}(Given
Name)\hspace{30mm}(Middle Name)

Institution: \rule[-1mm]{135mm}{.1mm}

Department: \rule[-1mm]{133mm}{.1mm}

Address: \rule[-1mm]{140mm}{.1mm}

\hspace*{15mm} \rule[-1mm]{140mm}{.1mm}

Fax: \rule[-1mm]{66mm}{.1mm}\ \ Phone: \rule[-1mm]{66mm}{.1mm}

Names of Accompanying Person(s): \hrulefill

{\bf Hotel Accommodations}:

\begin{tabular}{|l|c|lc|c|}\hline
Name of Hotel & Number of Rooms & \multicolumn{3}{|c|}{Period of
Stay}\\ \hline
1st Choice    & & IN & Sep. \rule[-1mm]{15mm}{.1mm} & \\
\hspace*{15mm}\rule{40mm}{.1mm} & \rule{15mm}{.1mm} Twin 
	& & Oct. \rule[-1mm]{15mm}{.1mm} & \\
1st Choice    & & OUT & Sep. \rule[-1mm]{15mm}{.1mm} &
\rule{10mm}{.1mm} Nights\\
\hspace*{15mm}\rule{40mm}{.1mm} & \rule{15mm}{.1mm} Single
	& & Oct. \rule[-1mm]{15mm}{.1mm} & \\ \hline
\end{tabular}

{\bf Payment}:

$\Box$ \ I would like to pay the above hotel deposit by Credit Card.

\hspace*{10mm}
$\Box$VISA\ \ $\Box$American Express\ \ $\Box$Diners\ \ $\Box$Master
(\rule[-1mm]{40mm}{.1mm})

\hspace*{10mm}
Card Number: \rule[-1mm]{60mm}{.1mm} Exp.\ Date: \rule[-1mm]{40mm}{.1mm} 

\hspace*{10mm}
Card Member: \rule[-1mm]{60mm}{.1mm} (in block letters)

\hspace*{10mm}
Card Member's Signature: \rule[-1mm]{60mm}{.1mm}

$\Box$ \ I enclose here with a bank draft covering the above hotel.\\
\hspace*{20mm}(Personal check will not be accepted)

\vspace{5mm}
Date: \rule[-1mm]{60mm}{.1mm}\ \  Signature: \rule[-1mm]{60mm}{.1mm}

\end{document}


