\documentstyle[12pt]{article}
\pagestyle{plain}
\title{LDOCE Error Guide \\ Version 1.0}
\author{Robert Krovetz\\
Department of Computer Science\\
University of Massachusetts, Amherst, MA 01003 \\
Krovetz@cs.umass.edu
}

\date{}
\begin{document}
\maketitle
\section{Introduction}
This is a guide to the errors I have found in the machine-readable version of the
Longman Dictionary of Contemporary English (LDOCE).  The guide refers to the first 
edition of the dictionary, and in particular, to the `lispified' version created by
Bran Boguraev.  If anyone is aware of other errors that are found in the
tape for the second edition, I would be glad to include those as well.

The guide is broken down into translation errors (errors that resulted from the 
conversion of the original tape into Lisp s-expressions), and errors in particular
fields: part-of-speech, subject codes, selectional restrictions, definitions,
and run-ons.   I have not examined the pronunciation field or the
grammar codes for the entire entry or specific senses.  I also did not attempt
to identify errors associated with the box codes on morphology.

If the reader is aware of
any errors that are not mentioned here, please let me know and they will be
incorporated into future versions.


\section{Translation Errors}
These errors are due to problems with parentheses, which
resulted in missed information when the tape was translated
into Lisp s-expressions.

Thanks to Robert Amsler for identifying these errors!

Note: Those marked with * are errors in the printed book;
Those marked with + are errors on the machine-readable typesetting tape
which were carried over into errors in the LISP version.
Those unmarked are errors in the LISP representation of legitimate
dictionary and typesetting tape text.

The escape character referred to below is the one used on the original tape, 
which was produced using Cambridge Lisp.  If you are using Common	
Lisp, you need to substitute the Common Lisp escape character (a backslash ($\backslash$))
for the percent sign.  The corrections I provide later on in this guide
use the Common Lisp escape character.

\begin{itemize}

\item aren't (usage note uses `)'s to enumerate `1)', `2)', `3)' resulting
	in lost text at end; patch=add a LISP escape character (`\%') to those `)'s
	and type in missing text from printed book.)

\item *barcarole (superfluous `)' after `Venice' resulting in premature closure
	of definition (8) text; patch=remove superfluous `)' and
	move to end of (8) text to reclose definition.)

\item *champ, v. (1st homograph) (superfluous `)' after BIT (1) in 1st sense
	of champ causing premature close of definition (8) text;
	patch=remove superfluous `)' and move to end of (8) to
	reclose definition.)

\item +engine (missing `(' in machine-readable tape before `from steam \%,'...;
	patch=add missing `(' to file).  [Note: there are also two missing senses
        from the entry - see comments in the section on definition errors].

\item *+foregoing, adj. (damaged 06 line missing field-delimiter ($<$) between
	`06' and `Wa5' (beginning of first field on original data tape).
	Probably resulting in omitted `Wa5' code from printed book
	(i.e. no comparative/superlative of adjective form); patch=add
	`(6' back and separate `Wa5' as first element of its record)

\item *great, n., 2nd homograph (superfluous `)' after `important people'
	in definition (8); patch=delete extra `)' and added closing
	`)' at end of entry.)

\item +misfire, n., 2nd homograph (empty pronunciation (3) record;patch=delete
	empty pronunciation record from entry.)

\item *pick up, v., adv. (superfluous closing parenthesis after `KNITing'
	`(*CA KNIT *CB ting' and before the sense numbers associated
	with `KNITing'; patch=delete `)' and add balancing `)' at
	end of entry).  [Note: there is also a missing sense from the entry -
        see comments in the section on definition errors].

\item *principled, adj. (superfluous `)' after `*CA PRINCIPLES *CB (3)';
	patch=remove extra `)' and add balancing `)' at end of entry.)

\item *rank, adj., 1st homograph, 2nd sense (missing `(' in grammatical code 
	of `B *46 with *44)' and loss of entire entry after
	2nd sense; patch=add `(' and manually add back missing
	rest of entry).  [Note: see the section on definition errors for
        the text of the missing portions].

\item *sail, v, 2nd homograph, sense 1b (superfluous `)' after `on the water'
	causing loss of rest of entry;patch=delete superfluous `)'
	and manually translate rest of entry (a lot!)).  [Note: see the
        section on definition errors for the text of the missing portions].
\end{itemize}



\section{Part of Speech errors}
Several words have font codes in their part of speech field.  These
are: {\it dalai  lama}, {\it defence}, {\it dentist}, {\it desalinize},
{\it diabolic}, {\it many}, {\it ochre}, {\it passion play}, {\it puffed}, 
{\it vibro-}, and {\it -intentioned}.

A number of words are missing the part of speech field entirely (i.e., the
s-expression for the part-of-speech field starts with a `5', and this
is missing from the entry).  There are 78 instances of this, and they are: 

\bigskip
\noindent
{\it aa}           -  X, second homograph   \\
{\it assault and battery} -  N  \\
{\it blowpipe}     -  N  \\
{\it boned}        -  ADJ  \\
{\it -built}       -  COMB. FORM, info given under field 7 \\
{\it callipers}    -  N  \\
{\it caveat emptor} - ? \\
{\it cent}         -  X, second homograph, period? \\
{\it chieftainship} - N \\
{\it clip}         -  N, fourth homograph \\
{\it cm}           -  X \\
{\it c}            -  X, third homograph \\
{\it C of E}       -  X \\
{\it common market} - N \\
{\it cu.}          -  X \\
{\it day return}   -  N \\			
{\it demagoguery}  -  N \\
{\it demobilize}   -  V \\
{\it demoralize}   -  V \\
{\it de-escalate}  -  V \\
{\it dewpond}      -  N \\
{\it dna}          -  N  (also abbreviation; X?) \\
{\it eating-house} -  N \\
{\it eminently}    -  ADV \\
{\it english horn} -  N \\
{\it escapology}   -  N \\
{\it ethnically}   -  ADV \\
{\it fagged}       -  ADJ \\
{\it fancies}      -  N \\
{\it father}       -  N, homograph 1 (priest) \\
{\it fatigues}     -  N \\
{\it favourite son} - N  (note spelling problem; only British spelling given) \\
{\it faze}         -  V \\
{\it feedbag}      -  N \\
{\it figured bass} -  N \\
{\it filet}        -  N \\
{\it fines herbes} -  N  (note spelling) \\
{\it fink}         -  N \\
{\it firstly}      -  ADV \\
{\it flamenco}     -  N \\
{\it flares}       -  N (a type of pants; should have P subcode?) \\
{\it F.O.B.}       -  X \\
{\it golden wedding} - N \\
{\it hammer}       -  N   \\			
{\it her}          - DETERMINER \\
{\it hole}         - V, second homograph \\
{\it instructor}   - N \\
{\it interrelate}  - V \\
{\it K.C.}         -  N (similar to Q.C.) \\
{\it liftboy}      - N \\
{\it lord's prayer} - N \\
{\it lyre}         - N \\
{\it mo}           - X (medical officer, modus operandi) \\
{\it multiplication table} - N \\
{\it ogre}         - N \\
{\it onerous}      - ADJ \\
{\it overburden}   - V \\
{\it pencil}       - N, 1st homograph \\
{\it pitchfork}    - V, 2nd homograph \\
{\it pompom}       - N \\
{\it pools}        - N \\
{\it prevention}   - N \\
{\it P T}          - X \\
{\it ram}          - N, 3rd homograph \\
{\it ranee}        - N \\
{\it rare earth}   - N \\
{\it reading desk} - N \\
{\it sealyham}     - N \\
{\it selective service} - N \\
{\it shinto}       - N \\
{\it stoker}       - N \\
{\it tablet}       - N \\
{\it trustworthy}  - ADJ \\
{\it trusty}       - N, 2nd homograph \\
{\it USSR}         - X (in analogy to US; but should also be N, I think). \\
{\it V.}           -  X \\

\bigskip
\noindent
The words below {\em do}  have field 5, but nothing before the field delimiter ($<$):

\medskip
\noindent
{\it cap.}        -  X  \\
{\it trapezium}   -  N \\
{\it bonsai}      -  N \\
{\it shovel}      -  V, 2nd homograph \\



Sometimes words are listed as having only an X for part-of-speech (e.g., US for
United States), but they also can be considered as N's.  The dictionary
also contains words that are certainly abbreviations (ESQ.), but have
N for their part-of-speech, so this is an inconsistency.

There are two entries, {\it by and large}, and {\it sopping}, that contain usage information
prior to the field delimiter;  this information 
 should come {\em after} the field delimiter.  In the word {\it both},  
there is a comma missing between the parts {\it determiner} and {\it predeterminer},
and in the word {\it bobsleigh} there should be a comma between parts V and N
(probably should be two separate definitions in analogy with {\it toboggan}).
In the word {\it darn}, the third homograph contains V in parentheses (along
with several other parts of speech separated by commas), and 
the parentheses should be removed.

The entries for {\it Black Art}, and {\it Norman} have a subcategorization code stored 
as part ofthe part-of-speech field (field 5), but it should be stored under field 6.
The entries for  {\it cubic}, {\it double - declutch}, {\it draughtsman}, and {\it perm}
contain a variant reference, which should be under field 4.

The entry for {\it -itis} is classified as a noun, but it should be indicated as being
a suffix.  The entry for {\it -ator} contains information in the definition field that
belongs in field 6.

There is a discrepency between the printed dictionary and the machine-readable
version with respect to the part-of-speech for 
{\it womanly}.  This word is categorized as an ADJ in the printed dictionary
but as an ADV on the tape.



\section{Subject code errors}
The subject codes are a two to four letter code that indicate the subject area
of the sense.  Longman supplies a list of the translations of these codes, but
there are many codes that are not included.   Some of them are simple typos, such
as a substitution of the number 1 for an `L', or the number 6 for a `G'.  I determined
which codes were not found in the list supplied by Longman, and then collected the
words that appeared with that unknown code.  This gave me a reasonable guess as to
what the code represented.  Here is a list of the ones that are most common:

\begin{description}
\item [LN - G]    Linguistics (Grammar) 
\item [MI - W]    Military (Weapons) 
\item [MS - C]    Measures (Cooking) 
\item [LT - R]    Literature (Rhetoric) [perhaps confused with LN - R] 
\item [SI - B]    Science (Biology) [should be under MD instead of SI] 
\item [BW - B]    Beer [should be under BV (Beverages), instead of Bowling] 
\item [BW - W]    Wine [ditto] 

\medskip

\noindent
\item [KI]        Kinship 
\item [ZZ]        Continent 
\item [1-]        Polar 
\item [2-]        Plant-and-Animal-Name 
\item [3-]        Tropical 
\item [4-]        Sea-Creature 

\end{description}

There were many instances where I was unable to make a reasonable guess
because there were only a few instances of a word with that code.
There were also errors for particular words, like `isoceles 
triangle' having MU - G (Music), where it should have
been MH - G (Math), or {\it insect} having code OZ - C, where it should have
been OZ - E.  I didn't pay much attention to these errors.  


Here is a list of simple typos: \\
\smallskip

\noindent
60  Geology  (should have been GO) \\
6O  Geology  (ditto) \\
G0  Geology  (ditto) \\
6a  Games    (should have been GA) \\
6b  Gambling (should have been GB) \\
6y  Gymnastics (should have been GY) \\
A6  Agriculture (should have been AG) \\
E6  Engineering (should have been EG) \\
M1  Military (should have been MI) \\
H1  History (should have been HI) \\
R1  Religion (should have been RL) \\




\noindent
60 - GO (Geography)    [cut out, atlas, artic] \\
6O - ditto     [physical geography, attract, cave in, atoll, bill, bound, \\
                dust, wash, wandering, watercourse, dustbowl, waste, \\
                arctic circle, drain, borderline, water, dry land, wake, \\
                bore, bourn, bourne, drainage basin, dry, ben, area, \\
                intercontinental, interior, breaker, waterfall, wallow, \\
                arid, dell, dowsing rod, drainage, wander] \\
G0 - ditto      [narrows] \\
GL - ditto      [defile] \\
6A - GA (Games) [bounce, defence, defense, drop, bound, dribble, \\
                 drawn, padding pool, blind man's bluff, dominoes, \\
                 backgammon, draughtsman, double - check, wallop, \\
                 walk off with, draw, athlete, dealer, birdseed (?), \\
                 distance, away, box kite, walk away from, walkaway, \\
                 draughts, athletics, dogtrot, arena, declare, \\
                 doll's house] \\
6F - ditto           [nine, plus fours] \\
6B - GB (Gambling)   [place bet, draw for, drawing, place, draw, wager] \\
GP - ditto               [croupier] \\
6Y - GY (Gymnastics)     [bout, box, boxing] \\
16 - 1- (Polar)          [penguin] \\
M6 - MG (Mining)         [ingot] \\
Y6 - YR (African)        [quagga, hartebeest] \\
A6 - AG (Agriculture)    [dust, arbor, arbour, swath, dustbowl, bloom, \\
                          arid, husk, drainage] \\
E6 - EG (Engineering)    [insulation, waste pipe, air conditioning, \\
                          black box, aqueduct, arc welding, dredge up, \\
                          armature, insulate, dry battery, atmospherics, \\
                          asphalt, dredge, insulator, dredger, arc] \\


\noindent
11 - 1- (Polar)           [artic, fulmar, narwhal] \\
21 - 2- (Plant and Animal Names)   [birch, puffin] \\
20 - 2- (Plant and Animal Names)   [islander, bog asphodal, fir,  \\
                 red admiral, rhododendron, raven, adder, mountain ash] \\
41 - 4- (Sea Creature)    [salmon, salmon trout] \\
H1 - HI (History)         [dynasty, galley, reconstruction] \\
M1 - MI (Military)        [wound, world war, lock, worst, lock onto,  \\
                           long - range] \\
Q1 - QI (American Indian) [whippoorwill, hickory, gopher, moose, \\
                           prarie dog, powwow, skunk, cardinal, pone,  \\
                           grizzly bear, wapitit, paleface, elk,  \\
                           poison ivy, mountain lion, wampum, popcorn, \\
                           papoose, squaw, oriole, panther] \\
R1 - RL (Religion)        [devout, diabolic] \\
S1 - SI (Science)         [residue] \\
Y1 - YI (Mid-East)        [dromedery, oryx, punic, wadi, wady, asp, \\
                           moorish arch, couscous, ghoul, bedouin, \\
                           date, near east, genius, jerboa, genie] \\

\noindent
B2 - BZ (Business)        [executive] \\
2C - ZC                   [seahorse] \\
2E - ZE                   [on] \\
2T - ZT                   [electronics] \\

\noindent
3O - 30 (Tropical)        [rain forest] \\
3Q - 30 (Tropical)        [rubber tree] \\
0Z - OZ (Zoology)         [primate] \\
O2 - OZ (Zoology)         [eggshell] \\
ZO - OZ (Zoology)         [primate] \\

\noindent
EE - ZE (Engineering, Electrical)      [dissipate] \\
EZ - EC (Economics)       [bourgeois, dodge] \\
FG - FO (Food)            [honeydew melon] \\
FT - LT (Literature)      [criticism] \\
HX - XH (India)           [sitar] \\
LL - CL (Clothes), RL (Religion) [waist, congregation] \\
MO - MD (Medicine)        [harelip, amphibian, welt, genus,  \\
                           courtship, soft palate] \\
ME - MD (Medicine)        [death rattle, death rate, debility, \\
                           deficiency disease] \\
UD - MD (Medicine) [only one word mis-classified, `wellbeing'] \\
NV - NU (Numismatics)     [change, wad] \\
SF - SI (Science)         [bivalve, biochemistry] \\
TM - TH (Theatre)         [amusement park] \\
Y- - YR (African)         [civet cat, warthog, palm oil, baobab, \\
                           chimpanzee, zebra, impala, hippopotomus, \\
                           gorilla, giraffe, panther] \\
Y5 - YR (African)         [gnu, springbok, mamba] \\
Y9 - YR (African)         [high - life, pygmy] \\

\noindent
RC - RLZC (Christianity)  [transubstatiation] \\
PJ - PTZJ (Journalism)    [subeditor] \\


Finally, the subject code field for the first homograph for {\it sting}
does not contain a subject code, but has (-*49-FOR-*44-)  [bizarre!].
This turned out to belong to a grammar code in the previous field.





\section{Selectional Restrictions}
The following words have a 0 (zero) in the selectional restriction field for 
one of their senses, and it should be changed to O: {\it bother, bugger,
drat, impale, neat, oppress, outsider, raise, recollect,} and {\it swarm
with}.  Those are the only occurrences of codes that are not on the
list given by Longman.





\section{Definition Errors}
This section describes errors encountered in the definitions in the dictionary.
This includes missed senses due to problems with parentheses, and discrepencies
between the printed dictionary and the machine-readable version.  I am not 
sure about the corrections I made for field 7 (which contains information
about subject codes and box codes), or about the correctness of the font code
information, but this is my best guess.  There are also several words that
have incorrect numerical information in their definitions (e.g., {\em millenium}
is defined as `(a period of 1,0 years)').  In all of these cases, a sequence
of three zeros has been replaced with a single zero.  Besides {\em millenium}
I have noticed this with {\em megadeath}, {\em megaton}, {\em million}, 
{\em millionaire}, {\em myriad}, {\em thousand}, and {\em grand}.  

\begin{description}

\item [alliance] Missing a right parenthesis at the end of the first definition.
                 Fix: add it, and eliminate the one at the end of the homograph.

\item [aquaplane] Missing a period following `esp' in the second sense of
                 the second homograph.

\item [andante] There is an extra set of parentheses before the initial parenthetical
                (a piece of music) and after `played'.  This is also an error in the
                printed dictionary.  Fix: remove the extra parentheses.

\item [antarctic circle] The definition has `line of latitude' in parentheses.  It
                seems like it should have `of latitude' in parentheses instead 
                (or else duplicate the word `line' before the parentheses so that
                 it is the same as `arctic circle').  Also, the parenthesis before
                 `South Pole' should close after the sense number `(1)'.


\item [bully]  Missing an escape character following the word `person' in the first homograph
      (a percent sign in the original tape; a backslash for the Common Lisp version).

\item [camp out] Missing a period following `esp' in the second sense.

\item [chill] Missing `to )' following `to (cause ' in sense 3 of homograph 1.
              This results in the other senses being grouped as part of that sense.
              Fix: add the missing text.


\item [engine] A parenthesis error caused the ommission of the following two senses:

\indent
(7 200 $\backslash< \backslash<$ RR-- $\backslash<$ ----J) \\
\indent
(8 also *45 locomotive *44 *63 a machine which pulls a \\
\indent
railway train) \\
\indent
(7 300 $\backslash< \backslash<$ VH-- $\backslash<$ ----J) \\
\indent
(8 *CA Fire Engine)



\item [eyrie]  The word `eyry' is mentioned as a variant in the headword, 
but it appears in the list twice.  This is also an error in the printed dictionary.


\item [favourite son]  The headword is spelled only according to the British spelling
(but has a language tag indicating it is American English!).


\item [firstly] The part of speech is incorrectly placed in the definition field
(resulting in its appearance in the list in Section 3).


\item [for] Missing a closing parenthesis in sense 26 of the first homograph.  This
causes senses 27-32 and the usage note to be considered part of that sense.  This
is a case in which the tape differs from the printed dictionary (which has the
missing parenthesis).  Fix: add a parenthesis following `(= in order that it should
begin)', and remove one at the end of the homograph.


\item [forebear]  The headword contains a hyphen after `forebear' as well as in the
embedded variant `forbear'.  The first hyphen should be omitted.


\item [Giant Panda] There is a period in the headword after the letter `P'.
  The problem is the use of a period instead of a bullet separator
(hex code *80) in field 1.   Patch: remove the period from the headword
and field 1 and substitute *80 for the period in field 1.


\item [go ahead]  Same problem as Giant Panda (in this case the period appears
after the `a').    


\item[have] The fourth sense of the second homograph is missing a closing
parenthesis (this is also an error in the printed dictionary).  Fix: add
the missing parenthesis after the word `bad', and remove one at the end
of the homograph.

\item [man] The dialect note is written as `Am' instead of `AmE'.


\item [nativity] The definition contains the usage information `(often cap. N)',
      and the `N' should be omitted.

\item [parameter] The entire definition for this word was omitted from the tape
(it {\em is} included in the printed version of the dictionary).

\item [party-coloured]   No variant is mentioned for `party - colored'.
The word isn't tagged as being British English, so `party - colored' should
be an acceptable variant (I haven't come across the word in American English
though; the error might just be an omission of a language indicator).


\item [pick up] A parenthesis error caused the ommission of the following sense: \\
\indent
(7 2100 $\backslash< \backslash< ---- \backslash<$ --I) \\
\indent
(8 *45 pick up and leave *46 infml *44 to collect $|$one's$|$ \\
belongings and leave suddenly))


\item [rank] A parenthesis error caused the omission of all senses in the first homograph following the
second sense.  Here is the entire entry:

\indent (((rank) \\
\indent (1 R0019600 $\backslash <$ rank) \\
\indent (2 1 $\backslash< \backslash<$)  \\
\indent (3 r *67 nk) \\
\indent (5 adj $\backslash <$) \\
\indent (6 Wa1 $\backslash <$) \\
\indent (7 100 $\backslash <$ B $\backslash < ---- \backslash < ----$ P) \\
\indent (8 (of a plant) too thick and widespread $|:|$ *46 cut the rank grass) \\
\indent (7 200 $\backslash <$ B (*46 with *44) $\backslash < ---- \backslash < ----$ N $ ---$ Y) \\
\indent (8 (of land) thickly covered (with useless plants)  \\
\indent $|:|$ *46 rank soil.  The field was rank with *ca WEEDS *CB *44 S)  \\
\indent (7 300 $\backslash <$ B $\backslash < ---- \backslash < ----$ P)  \\
\indent (8 (of smell or taste) very strong and unpleasant
\indent $|:|$ *46 rank tobacco *44) \\
\indent (7 400 $\backslash <$ A $\backslash < ---- \backslash < ----$ H)  \\
\indent (8 having mistakenly the bad nature of what is named \\
\indent  $|:|$ *46 He$\backslash$'s a rank beginner at the job  $|,|$ but he $\backslash$'ll learn) \\
\indent (10 1 $\backslash < \backslash \backslash$  ly $\backslash < \backslash <$ adv $\backslash < \backslash <$)  \\
\indent (10 2 $\backslash < \backslash \backslash$ ness $ \backslash < \backslash <$ n $\backslash <$ U $\backslash <$)))


\item [ride] The entry has a comment at the end of homograph 1 - `see also, 
ride for a *CA FALL *CB *8B *44 (17)'.  This is not found in the printed version.
There is also no sense 17 for homograph 2 of the word {\it fall}, and I've
never heard of the expression before.  


\item [sail] A parenthesis error caused the ommission of most of the second homograph.  Here
is the entire entry:

\indent ((sail)  \\
\indent (1 S0009400 $\backslash <$ sail)  \\
\indent (2 2 $\backslash < \backslash <$)  \\
\indent (5 v $\backslash <$) \\
\indent (7 100 $\backslash <$ L $|;|$ X $|:|$ 9 $\backslash <$ NA-- $\backslash < ----$ 1 $----$ J) \\
\indent (8 *45 a *44 (of any ship) to travel on the water  \\
\indent $|:|$ *46 Watch the ships sail by *45 b *44 to command or direct (any ship) on the water  \\
\indent $|:|$ *46 The captain sailed his ship through the narrow passage) \\
\indent (7 200 $\backslash <$ L9 $\backslash <$ NA$-- \backslash < ----$H)  \\
\indent (8 (of people) to travel by ship $|:|$ *46 We sailed across the Atlantic in 5 days) \\
\indent (7 300 $\backslash <$ I0 $\backslash <$ NA $-- \backslash < ----$ J)  \\
\indent (8 to begin a voyage $|:|$ *46 We sail / our ship sails tomorrow (for New York))  \\
\indent (7 400 $\backslash <$ T1 $\backslash <$ NA $-- \backslash < ----$ J $----$ L)  \\
\indent (8 to move or travel on (water) 
\indent $|:|$ *46 One of the first men / ships to sail the Atlantic ocean)  \\
\indent (7 500 $\backslash <$ I0 $\backslash <$ NA $-- \backslash < ----$ H)  \\
\indent (8 to make short trips as a sport in a small boat \\
\indent  with sails (often in the phr $|.|$ *45 go sailing))  \\
\indent (7 600 $\backslash <$ L9 $\backslash <$ LNZF $\backslash < ----$ O)  \\
\indent (8 to move proudly $|,|$ smoothly $|,|$ or easily  \\
\indent $|:|$ *46 Sailed through the difficult examination with no 
\indent mistakes $\backslash$    Birds sailing across the sky 
\indent *63 see also *45 sail close to the *44 *CA WIND *CB *8A (3))) 


\item [Tues] This entry is missing in the printed version of the dictionary.  Both the printed
and machine-readable versions contain entries for  {\it Mon}, and {\it Fri}, but omit the entries
for {\it Sat, Sun, and  Thur}.  There is an entry for {\it Wed} (besides the verb sense), but it
differs from the others by omitting the period.  


\item [west] The word `is' is missing following the word `Reading' in the third example sentence
for the first sense of the first homograph.  This is another
difference between the printed copy and the machine-readable tape.


\end{description}








\section{Run-on errors}
A run-on is a word or suffix found at the end of an entry that indicates a
word form that is different from the headword.  They are indicated on the tape 
by an s-expression that begins with `10'.  

Some of the errors with this field are due to an extra hyphen, an embedded font code, or
to variantions in structure.  These are:

\smallskip

\noindent
{\it absorbing}      - field is present, but doesn't contain any run-on information \\
{\it agriculture}    - a parenthesized AL indicates optionality.  Needs to be split into two run-ons. \\
{\it cranial}        - unnecessary `-' at start of field \\
{\it criminal}       - unnecessary `-' at start of field \\
{\it fratricide}     - font code breaks up `-CIDAL' \\
{\it friendly}       - font code breaks up `-LINESS' \\
{\it hypodermic}     - unnecessary `-' at start of field \\
{\it hypothetical}   - unnecssary `-' at start of field \\
{\it phantasmagoria} - unnecessary `-' starting `phantasmagoric' and `phantasmagorical' \\
{\it photosensitize} - font code and hyphen break up `-TIZATION' \\


\medskip

Most of the errors concerned missing part-of-speech information in the
field entry.  These are:

\smallskip

\noindent
{\it accursed}    - missing ADV \\
{\it atrocious}   - missing ADV \\
{\it careful}     - missing N \\
{\it circumnavigate} - missing N \\
{\it corresponding} - missing ADV \\
{\it cursive}     - missing ADV \\
{\it cytology}    - missing N \\
{\it efficient}   - missing ADV \\
{\it egotist}     - missing ADJ \\
{\it eulogistic}  - missing ADV \\
{\it flaccid}     - missing N \\
{\it fleet}       - missing N \\
{\it government}  - missing ADJ \\
{\it governor}    - missing N \\
{\it hairy}       - missing N \\
{\it heated}      - missing ADV \\
{\it heptagon}    - missing ADJ \\
{\it hoard}       - missing N \\
{\it homogeneous} - missing N \\
{\it hooligan}    - missing N \\
{\it hoyden}      - missing ADJ \\
{\it hurried}     - missing ADV \\
{\it humpback}    - missing ADJ \\
{\it hypersensitive} - missing N \\
{\it inanimate}   - missing ADV \\
{\it inauspicious} - missing ADV \\
{\it joyous}      - missing ADV \\
{\it lithographic} - missing ADV \\
{\it pejorative}  - missing ADV \\
{\it questionable} - missing ADV \\
{\it rampage}     - missing ADJ \\
{\it rare}        - missing N \\
{\it retrench}    - missing N \\
{\it saint}       - missing N \\
{\it salivate}    - missing N \\
{\it salt}        - missing N \\
{\it sanguine}    - missing N \\
{\it scold}       - missing N \\
{\it scurrilous}  - missing N \\
{\it serious}     - missing N \\
{\it separable}   - missing N \\
{\it shirk}       - missing N \\
{\it smooch}      - missing ADJ \\
{\it surprising}  - missing ADV \\
{\it sweeping}    - missing ADV \\
{\it syllabify}   - missing N \\
{\it timeliness}  - missing ADV \\
{\it unclean}     - missing N \\

There are also a number of changes that have to be made to the run-ons themselves: 
(a few of these are due to the algorithm I used with the run-ons):

\smallskip

\noindent
{\it anachronism} : change `-TIC' and `-TICALLY' to `-ISTIC' and `-ISTICALLY' \\
{\it anemia}      : change `-C' and `-CALLY' to `-IC' and `-ICALLY' \\
{\it baleful}     : change `FULLY' to `-FULLY' \\
{\it bat}         : change `ER' to `TER' \\
{\it cannibalism} : change `-TIC' to `-ISTIC' \\
{\it catalepsy}   : change `-TIC' to `-EPTIC' \\
{\it determinism} : change `-TIC' to `-ISTIC' \\
{\it dive}        : change `ER' to `R' \\
{\it emigrate}    : change `-ION' to `-TION' \\
{\it entral}      : change `INGLY' to `-LLINGLY' \\
{\it euphoria}    : change `-CALLY' to `-ICALLY' \\
{\it existenialism} : change  `IST' to `-LIST' \\
{\it hedonism}    : change `ISITC' to `-ISITC' \\
{\it heterogeneous} : change `-ITY' to `-EITY' \\
{\it homogeneous} : change `-ITY' to `-EITY' \\
{\it hypnosis}    : change `-TIC' and `TICALLY' to `-OTIC' and `-OTICALLY' \\
{\it iamb}        : change `-IC' to `IC' \\
{\it lusty}       : change `ILY' and `INESS to `-ILY' and `-INESS' \\
{\it maladjusted} : change `-MENT' to `-TMENT' \\
{\it merrymaking} : change `-ER' to `-KER' \\
{\it numismatics} : change `-IST' to `-TIST' \\
{\it openly}      : change `-NESS' to `-NNESS' \\
{\it paternalism} : change `-TIC' and `-TICALLY' to `-ISTIC' and `-ISTICALLY' \\
{\it pharisaic}   : change `-ISM' to `-AISM' \\
{\it polytheism}  : change `-TIC' to `-ISTIC' \\
{\it psychokenesis} : change `-TIC' and `-TICALLY' to `-ETIC' and `-ETICALLY' \\
{\it pugalism}    : change `-TIC' to `-ISTIC' \\
{\it repatriate}  : change `-ATION' to `-IATION' \\
{\it simultaneous} : change `-ITY' to `-EITY' \\
{\it skate}       : change `ER' to `R' \\
{\it smutty}      : change `TILY' and `TINESS' to `-ILY' and `-INESS' \\
{\it speeding}    : change `-ER' to `-DER' \\
{\it wan}         : change `-LY' to `LY' \\
{\it windily}     : change `-NESS' to `-INESS' \\
{\it wrongdoing}  : change `-ER' to `-OER' \\



\section{Miscellaneous errors}
The tape I received from Longman contained the following messages:

\begin{itemize}

\item near the entry for {\it buhl} is a message `*** priority changed from 1.39
   to  1.45'  

\item within the definition for {\it bandy} is a message `*** priority changed
   from 84.9 to 139.5'

\end{itemize}

I don't know if the tapes given to other researchers also contained these
messages.

\section{Acknowledgments}
Thanks to Bob Amsler for supplying me with the errors described in the
translation section.

\end{document}


