From crabapple.srv.cs.cmu.edu!fs7.ece.cmu.edu!europa.eng.gtefsd.com!howland.reston.ans.net!spool.mu.edu!olivea!koriel!sh.wide!wnoc-tyo-news!virgil!ccgwy!soprano!wakatuki Wed Aug 18 13:08:44 EDT 1993
Article: 18419 of comp.ai
Xref: crabapple.srv.cs.cmu.edu comp.ai:18419
Newsgroups: comp.ai
Path: crabapple.srv.cs.cmu.edu!fs7.ece.cmu.edu!europa.eng.gtefsd.com!howland.reston.ans.net!spool.mu.edu!olivea!koriel!sh.wide!wnoc-tyo-news!virgil!ccgwy!soprano!wakatuki
From: wakatuki@ttl.cas.uec.ac.jp (Mituo WAKATUKI)
Subject: The 4th Workshop on Algorithmic Learning Theory (ALT'93)
Message-ID: <CBy39s.4Aq@voice.cas.uec.ac.jp>
Sender: news@voice.cas.uec.ac.jp
Nntp-Posting-Host: comet.ttl.cas.uec.ac.jp
Organization: The University of Electro-Communications, Tokyo, Japan.
Date: Wed, 18 Aug 1993 07:53:51 GMT
Lines: 493

-------


           **********************************************************

            The 4th Workshop on Algorithmic Learning Theory (ALT'93)

           **********************************************************


%%%%%%%%%%%%%%%%%%%%% << Final program in LaTeX format >> %%%%%%%%%%%%%%%%%%%%%%
% -------------------------- [ ALT'93 Final Program ] --------------------------
\documentstyle[11pt,twocolumn]{article}

\textheight=27cm
\textwidth=18cm
\topmargin=-2.5cm
\oddsidemargin=-12mm
\evensidemargin=\oddsidemargin
\columnsep=5mm

\newenvironment{session}[3]{\vspace{2mm}\begin{tabular}{l@{\hspace{2mm}}ll}#1 & \bf #2 & \bf #3\end{tabular}\par
\vspace{2mm}\begin{list}{}{\setlength{\leftmargin}{10mm}\setlength{\itemsep}{2mm}\setlength{\topsep}{-2mm}}}{\end{list}\vspace{2mm}}
\newcommand{\talk}[2]{\baselineskip=6pt
\item[] {\scriptsize\bf #1} \\[3pt]
{\em #2}}
\newcommand{\talkA}[1]{\baselineskip=6pt
\item[] {\scriptsize\bf #1}}
\newcommand{\chair}[1]{\vspace{-1.5mm}\begin{flushright}\scriptsize Chair : {#1}\end{flushright}\vspace{-.5mm}}
\newcommand{\BREAK}[2]{\vspace{2mm}\begin{tabular}{ll}#1 & \bf #2\end{tabular}\par\vspace{1mm}}


\pagestyle{empty}

\begin{document}
\begin{center}
  {\bf Fourth International Workshop on} \\
  {\bf Algorithmic Learning Theory} \\[2mm]
  {\sf ALT'93} \\[4mm]
  {\large November 8 - 10, 1993} \\
  {\Large\bf  Program}
\end{center}

%\vspace{-10mm}
\baselineskip=15pt

It is with great pleasure that we invite researchers in all areas 
related to algorithmic and computational theories of learning 
to the {\em Fourth International Workshop on Algorithmic Learning Theory}
(ALT'93), which will be held in Tokyo, November 8-10, 1993.  
The ALT conferences have been held annually since 1990, 
and are organized and sponsored by the Japanese Society 
for Artificial Intelligence(JSAI).  
The main objective of this workshop is to 
provide an open forum for discussions and exchanges of ideas 
between researchers from various backgrounds 
in this emerging, inter-disciplinary field of research. 

By now the importance of `machine learning' to the 
success of the next generation AI systems 
has been widely recognized and accepted. 
At the same time, decades of theoretical research in `inductive inference' 
and its complexity theoretic analogue have led to the emergence of 
`algorithmic/computational learning theory.'
It is our goal and our hope to bring together researchers from both 
of these fields, and all related areas, 
to have mutually beneficial research interactions. 
This year we are fortunate to have three invited 
lectures by distinguished researchers: 
 ``Identifying and Using Patterns in Sequential Data(by Dr.\ P.\ Laird, 
NASA Ames Research Center)'',  
 ``Learning Theory Toward Genome Informatics(by Prof.\ S.\ Miyano, 
Univ.\ of Kyushu)'' and ``Optimal Layered Learning : A PAC approach to
incremental sampling(by Prof.\ S.\ Muggleton, Oxford Univ.)''.   

The conference site,  the University of Electro-Communications, 
is located in Chofu city, just 10 miles west of the center of Tokyo,  
which provides an ideal base to visit sites in Tokyo.   
We hope that the participants  will have enjoyable 
time both inside  and outside the conference site.
\newpage


\noindent{\bf November 8 (Monday)}
\footnotesize

\BREAK{12:00-13:00}{Registration}

\begin{session}{13:00-13:10}{Opening Address}{}
\talk
{S. Kobayashi (Tokyo Inst. of Tech.,Conf. Chair)}{}
\end{session}

\begin{session}{13:10-14:10}{Session 1}{Invited Talk}
\chair{K. Furukawa (Keio Univ.)}
\talk
{S. Muggleton (Oxford Univ.)}
{Optimal Layered Learning : a PAC approach \\ to incremental sampling}
\end{session}

\BREAK{14:10-14:30}{Break}

\begin{session}{14:30-15:30}{Session 2}{Inductive Logic}
\chair{H. Motoda (Hitachi)}
\talk
{M. Hagiya (Univ. of Tokyo)}
{A Typed $\lambda$-Calculus for Proving-by-Example and \\ Bottom-Up Generalization Procedure}
\talk
{T.Sato and S.Akiba (ETL)}
{Inductive Resolution}
\talk
{J.Arima and H.Sawamura (Fujitsu Labs.)}
{Reformulation of Explanation by Linear Logic \\ --- Toward Logic for Explanation}
\end{session}

\BREAK{15:30-15:50}{Break}

\begin{session}{15:50-16:50}{Session 3}{Inductive Logic and Inference}
\chair{S. Muggleton (Oxford Univ.)}
\talk
{A. Yamamoto (Hokkaido Univ.)}
{Generalized Unification as Background \\ Knowledge in Learning Logic Program}
\talk
{J. Barzdins et al. (Univ. of Latvia)}
{Towards Efficient Inductive Synthesis \\ of Expressions from input/output Examples}
\talk
{K. P. Jantke and S. Lange (HTWK Leipzig)}
{Case-Based Representation and Learning \\ of Pattern Languages}
\end{session}

\BREAK{16:50-17:10}{Break}

\begin{session}{17:10-18:10}{Session 4}{Approximate Learning I}
\chair{H. Imai (Univ. of Tokyo)}
\talk
{J. R. Cuellar (Siemens) and H. U. Simon (Univ. of Dortmund)}
{Neural Discriminant Analysis}
\talk
{M. Iwayama, N. Indurkhya and H. Motoda (Hitachi)}
{A New Algorithm for Automatic \\ Configuration of Hidden Markov Models}
\talk
{A. Sakurai (Hitachi)}
{On the VC-dimension of Depth Four Threshold \\ 
Circuits and the Comlexity of Boolean-valued Functions}
\end{session}

\BREAK{18:10-18:30}{Break}

\BREAK{18:30-20:00}{\em RECEPTION}
%\newpage

\vspace{5mm}
\noindent{\normalsize\bf November 9 (Tuesday)}

\BREAK{~9:00-~9:30}{Registration}

\begin{session}{~9:30-10:30}{Session 5}{Query Learning}
\chair{O. Watanabe (Tokyo Inst. of Tech.)}
\talk
{A. Nakamura and N. Abe (NEC)}
{Exact Learning of Linear Combinations of \\ Monotone Terms from Function Value Queries}
\talk
{R. Siromoney et al. (Madras Christian College)}
{Thue Systems and DNA --- A Learning \\ Algorithm for A Subclass}
\talk
{Y. Ishigami and S. Tani (Waseda Univ.)}
{The VC-dimension of Finite Automata with $n$ States}
\end{session}

\BREAK{10:30-10:50}{Break}

\begin{session}{10:50-11:50}{Session 6}{Inductive Inference I}
\chair{K. P. Jantke (HTWK Leipzig)}
\talk
{Y. Mukouchi and S. Arikawa (Kyushu Univ.)}
{Inductive Inference Machines That \\
 Can Refute Hypothesis Space}
\talk
{R. Freivalds (Univ. of Latvia) and C. H. Smith (Univ. of Maryland)}
{On the Duality between Mechanistic \\
 Learners and What it is They Learn}
\talk
{S. Jain (National Univ. of Singapore) and A. Sharma (Univ. of New South Wales)}
{On Aggregating Teams of Learning Machines}
\end{session}

\BREAK{11:50-13:00}{LUNCH}

\begin{session}{13:00-14:00}{Session 7}{Invited Talk}
\chair{T. Sato (ETL)}
\talk
{P. Laird  (NASA Ames Res. Cent.)}
{Identifying and Using Patterns in Sequential Data}
\end{session}

\BREAK{14:00-14:20}{Break}

\begin{session}{14:20-15:20}{Session 8}{EBL}
\chair{M. Numao (Tokyo Inst. of Tech.)}
\talk
{K. Yoshida, H. Motoda, and N. Indurkhya (Hitachi)}
{Integration of Learning Methods by Colored Digraphs}
\talk
{M. Suwa and H. Motoda (Hitachi)}
{A Perceptual Criterion for Visually \\ Controlling Learning}
\talk
{S. Kobayashi (Univ. of Electro-Communications)}
{Learning Strategies using Decision Lists}
\end{session}

\BREAK{15:20-15:40}{Break}

\begin{session}{15:40-16:20}{Session 9}{Inductive Inference II}
\chair{C. H. Smith (Univ. of Maryland)}
\talk
{J.Viksna (Univ. of Delaware)}
{Learning with Growing Quality}
\talk
{R. Daley and B. Kalyanasundaram (Univ. of Pittsburgh)}
{Use of Reduction Arguments on Detertabular}
\end{session}

\BREAK{16:20-17:00}{Break}

\vspace{3mm}
\begin{tabular}{llp{50mm}}
17:00-18:30 & \bf Session 10 & \bf Impromptu Talks
\end{tabular}
\par\vspace{-1mm}
\begin{flushright}\scriptsize
Chair : Naoki Abe (NEC)
\end{flushright}
%\newpage

\vspace{5mm}
\noindent{\normalsize\bf November 10 (Wednesday)}

\BREAK{~9:00-~9:30}{Registration}

\begin{session}{9:30-10:30}{Session 11}{Approximate Learning II}
\chair{S. Miyano (Kyushu Univ.)}
\talk
{E. Takimoto and A. Maruoka (Tohoku Univ.)}
{On the Sample Complexity of Consistent \\
 Learning with One-Side Error}
\talk
{A. Shinohara (Kyushu Univ.)}
{Complexity of Computing Vapnik-Chervonenkis \\ Dimension}
\talk
{S. Hasegawa, H. Imai and M. Ishiguro (Univ. of Tokyo)}
{$\epsilon$-approximations of $k$-label spaces}
\end{session}

\BREAK{10:30-10:50}{Break}

\begin{session}{10:50-11:50}{Session 12}{Inductive Inference III}
\chair{R. Freivalds (Univ. of Latvia)}
\talk
{T. Moriyama and M. Sato (Univ. of Osaka Pref.)}
{Properties of Language Classes with \\
 Finite Elasticity}
\talk
{S. Kapur (Univ. of Pennsylvania)}
{Uniform Characterizations of Various \\
 Kinds of Language Learning}
\talk
{T. Knuutila (Univ. of Turku)}
{How to Invent Characterizable Inference \\
 Mothod for Regular Languages}
\end{session}

\BREAK{11:50-13:00}{LUNCH}

\begin{session}{13:00-14:00}{Session 13}{Invited Talk}
\chair{Y. Sakakibara (Fujitsu Labs.)}
\talk
{S. Miyano (Kyushu Univ.)}
{Learning Theory Toward Genome Informatics}
\end{session}

\BREAK{14:00-14:20}{Break}

\begin{session}{14:20-15:20}{Session 14}{New Learning Paradigms}
\chair{T. Yokomori (Univ. of Electro-Communications)}
\talk
{N. Zhong and S. Ohsuga (Univ. of Tokyo)}
{A Decomposition Based Induction Model for \\
 Discovering Concept Clusters from Databases}
\talk
{J.-G.Ganascia (Univ. of Paris)}
{Algebraic Structure of Some Learning Systems}
\talk
{S. Tsumoto and H. Tanaka (Tokyo Medical and Dental Univ.)}
{Induction of Probabilistic Rules  \\
 Based on Rough Set Theory}
\end{session}

\begin{session}{15:20-15:30}{Closing Address}{}
\talkA
{E. Tomita (Univ. of Electro-Communications, Local Arrang. Chair)}
\end{session}

\BREAK{}{Adjorn}

\end{document}
% ------------------------------ (End of Program) ------------------------------


%%%%%%%% << Registration & Hotel Accommodation Form in LaTeX format >> %%%%%%%%%
\documentstyle[fleqn]{article}

\setlength{\topmargin}{-10.0mm}
\setlength{\oddsidemargin}{0pt}
\setlength{\evensidemargin}{0pt}
\setlength{\headheight}{0pt}
\setlength{\headsep}{12pt}
\setlength{\footheight}{12pt}
\setlength{\footskip}{30pt}
\setlength{\textheight}{720pt}
\setlength{\textwidth}{530pt}
\setlength{\columnsep}{10pt}
\setlength{\columnseprule}{0pt}
\setlength{\hoffset}{-15.0mm}
\setlength{\voffset}{3.0mm}
\renewcommand{\arraystretch}{0.9}
\pagestyle{empty}

\begin{document}

\begin{small}
\twocolumn[%
\begin{center}
{\large {\bf ALT'93 (November 8-10, 1993)}}\\
{\large {\bf Fourth Workshop on Algorithmic Learning Theory}}\vspace{5.0mm}\\
{\large {\bf REGISTRATION AND ACCOMMODATION FORM}}\vspace{3.0mm}\\
{\small Please complete (block letters only) appropriately
and return this form to:}\vspace{3.0mm}\\
\begin{tabular}{ll}
 & {\tt Prof. Etsuji Tomita, Local Arrangement Chair of ALT'93,} \\
 & {\tt Dept. of Communications and Systems Engineering,} \\
{\bf Early registration :}
 & {\tt Univ. of Electro-Communications,} \\
For payments received
 & {\tt 1-5-1 Chofugaoka, Chofu, Tokyo 182, JAPAN} \\
before \( \underline{\mbox{September 30, 1993}} \).
 & {\tt (Phone) +81-424-83-2161 ex.4375} \\
 & {\tt (Fax) +81-424-82-3055} \\
 & {\tt alt93@cs.uec.ac.jp }
\end{tabular}
\end{center}]
\vspace{5mm}
\noindent
\hspace*{1mm}
Title: \hspace{10mm} $\Box$Prof./$\Box$Dr./$\Box$Mr./$\Box$Ms.\vspace{3mm}\\
\noindent
\begin{tabular}{llll}
             &                 &                &                  \\
Name:        & \hspace*{19mm}  & \hspace*{19mm} & \hspace*{19mm}   \\
\cline{2-4}
             & {\footnotesize Family} & {\footnotesize First} & {\footnotesize Middle}  \\
             &                 &                &                  \\
             &                 &                &                  \\
Affiliation: &                 &                &                  \\
\cline{2-4}
             &                 &                &                  \\
             &                 &                &                  \\
Address:     &                 &                &                  \\
\cline{2-4}
             & {\footnotesize Street} &                & {\footnotesize City}    \\
             &                 &                &                  \\
             &                 &                &                  \\
\cline{2-4}
             & {\footnotesize State}  & {\footnotesize Zip}   & {\footnotesize Country} \\
             &                 &                &                  \\
\end{tabular}
\vspace{1.0mm}\\
\hspace*{1mm}
Telephone: \underline{\hspace{29mm}} \hfill Telex: \underline{\hspace{29mm}}\vspace{5mm}\\
\hspace*{1mm}
Facsimile: \underline{\hspace{29mm}} \hfill E-mail: \underline{\hspace{29mm}}\vspace{3mm}\\
\noindent
{\normalsize {\bf Registration Fee}} (includes proceedings and a reception party)
\begin{description}
\item [$\Box$] 25,000 {\tt Yen} / 22,000 {\tt Yen}$^*$ for a member
of JSAI and cooperative\linebreak[4]
associations\par
\hfill \underline{\hspace{30mm}} {\tt Yen}
\item [$\Box$] 28,000 {\tt Yen} / 25,000 {\tt Yen}$^*$ for others\par
\hfill \underline{\hspace{30mm}} {\tt Yen}
\item [$\Box$] 12,000 {\tt Yen} / 11,000 {\tt Yen}$^*$ for a fulltime student\par
\hfill \underline{\hspace{30mm}} {\tt Yen}\\
\hfill {\footnotesize $^*$ For payments received \underline{before} September 30, 1993}
\item [$\Box$] 10,000 {\tt Yen} for Extra Banquet Ticket
$\times$ \underline{\hspace{8mm}} person(s)\par
\hfill  \underline{\hspace{30mm}} {\tt Yen}
\end{description}
\hspace{25mm} $\rhd$ {\bf Total amount:}
\hfill  \underline{\hspace{30mm}} {\tt Yen}\vspace{3mm}\\
\noindent
{\normalsize {\bf Payment}}
\begin{description}
\item [$\Box$] Bank Transfer (No personal checks accepted)\\
I have sent the above amount on \underline{{\footnotesize (Date)} \hspace{20mm}}\\
\noindent
through \underline{{\footnotesize (Name of Bank)} \hspace{43mm}}\\
\noindent
to the following account in {\bf Japanese Yen.}\vspace{-1.0mm}
\begin{center}
Bank: {\tt DAI-ICHI-KANGYO BANK, Iidabashi Branch}\\
Account: {\tt ALT93 Jinkou-Chinou-Gakkai}\\
Account No. {\tt 061 - 1773862}\vspace{-1.0mm}
\end{center}
\item [$\Box$] Payment At the Conference Desk (By cash only)\\
I am sure that I will pay the above amount at the conference desk of
ALT'93.
\end{description}
\vspace{1mm}
{\small
\noindent
Notes:
\begin{enumerate}
\item In case of Bank Transfer, participant's name and ``{\tt ALT93
Jinkou-Chinou-Gakkai}'' should be specified on payment. (If a group
payment has been made, please indicate the name of the remitter and
attach a list of participants to confirm accurately their individual
payments.)
\item In case of Bank Transfer, the handling charge  
should be added to the total amount and proof of payment should accompany this form.
\item In case of Payment At the Conference Desk, the date when this form is received
is treated as the date when payments are received.
\end{enumerate}
}
%%
\vspace{3mm}
\noindent
{\normalsize {\bf Hotel Accommodation}}\\
\noindent
Rooms for participants will be booked by local arrangement\linebreak[4]
committee (booked first-in-first-served).\vspace{2.0mm}\\
Hotels: \hspace{3mm} {\tt Creston Hotel},\\
\hspace*{15.8mm}{\tt Hotel Twins Chofu}\\
\hspace*{14.0mm}(Approx. 10 minutes walk to the conference site)\vspace{2.0mm}\\
Creston Hotel
\begin{description}
\item [$\Box$] 9,700 {\tt Yen} for a single room per night
\item [$\Box$] 13,500 {\tt Yen} for a double room per night
\end{description}
Twins Chofu
\begin{description}
\item [$\Box$] 8,000 {\tt Yen} for a single room per night
\end{description}
Date of arrival: \hspace{3mm} Nov. \underline{\hspace{10mm}} \hspace{5mm}\\
Date of departure: Nov. \underline{\hspace{10mm}} $=$ \underline{\hspace{10mm}} night(s)\\
\noindent
{\small
Notes: 
\begin{enumerate}
\item No meal is included in the above room rates.
\item 3 \% $\sim$ 6 \% tax and 10\% service charge will be added to
the bill when checking out.
\item Only night(s) on Nov.7-10 will be booked, and especially for double rooms on Nov.6-10.
\item Alternative selection :
Town Hotel Chofu (30 minutes walk to the conference site, Phone +81-3-3454-6376).
\end{enumerate}
}
\vspace{3mm}
\noindent
{\normalsize {\bf Cancellation}}\\
\noindent
I agree that I will pay the following fee in case of cancellation.\\
1. Registration:\hspace{15mm}\dotfill 30\% of registration fee\\
2. Accomodations:\par
\hspace{7mm} No notice given before the first night of stay\par
\hspace{35mm} \dotfill 100\% of room rate for 1 night
\vfill
\noindent
Date: \hspace{6mm} \underline{\hspace{20mm}} \hspace{5mm}\vspace{3mm}\\
\noindent
Signature: \underline{\hspace{70mm}}
\end{small}
\end{document}
% -------------- (End of Registration & Hotel Accommodation Form) --------------

-------                             Written: 1993.8.18 16:47
**  Department of Communications and Systems Engineering, **
 **          Faculty of Electro-Communications,          **
  **      The University of Electro-Communications,     **
   **     1-5-1 Chofugaoka, Chofu, Tokyo, 182 JAPAN    **
    **        E-mail wakatuki@ttl.cas.uec.ac.jp       **
   **         Phone  +81-424-83-2161 (Ext.3243)        **
  **          Fax.   +81-424-89-6943                    **
 **                                                      **
**                    Mitsuo WAKATSUKI                    **


