\documentstyle[a4]{article}
\pagestyle{myheadings}
\markboth{Author:kk}{Doc.No. AID/newsletter/1}
\setlength{\textheight}{9in}
\setlength{\topmargin}{0in}
\setlength{\headheight}{.3in}
\setlength{\headsep}{.5in}
\parindent=0pt
\parskip= .35\baselineskip plus .0833333\baselineskip minus .0833333\baselineskip
\begin{document}
Artificial Intelligence for Development\\
Document No: AID/newsletter/1\\
Last Modified and by whom: 22/10/90 kk \\
Distributed on:  22/10/90\\
\LARGE
\begin{center} Artificial Intelligence for Development\\
Newsletter Number 1\\October 1990\\
\end{center}
\normalsize

This newsletter is being sent by both electronic and paper mail. If
you no longer wish to receive it, or have any contributions or
comments on its content, please contact KK by any of the normal
methods. If you would like a copy of the literature database (which is
in BibTeX format but perfectly comprehensible to ordinary folk) or
have anything to contribute to it, please mail ER. Addresses are at
the end of this document. 

\section{Next meetings:}
The next meeting will be on {\bf Thursday November 1st} at 5.30pm in
room F10 of the AI Department. Please arrive promptly so that we can
let you into the building. The meeting will take the following form: 
\begin{enumerate}
\item Matters outstanding from last meeting (workshop etc.)
\item Short presentation on {\em ``A Cost-Effective Computational Diagnostic
Aid for Zimbabwe Nurses''}
\item Discussion of presentation
\item Software demonstrations
\item Retire to a local hostelry (optional)
\end{enumerate}

On {\bf Thursday January 10th}
Robert Muetzelfeldt will speak on {\em ``A knowledge-based land use
allocation program - potential for development''} and give a software
demonstration. 

On {\bf Thursday March 7th}
Dan Walker will speak on {\em  ``The representation of indigenous
ecological knowledge about agroforestry''}.

On {\bf Thursday May 2nd}
Teresa Anderson will speak on {\em ``AI for Rural Electrification in Nepal:
Experiences with laptops, 14000 feet up in the Annapurnas''}  and give
a software demonstration. 

\section{News from around the world}
SUN MICROSYSTEMS SIGNS FIRST DISTRIBUTOR AGREEMENT IN AFRICA\\
(OK, I know it's a bit out of date, but I just thought I'd start this
section and see if anyone had any news...)\\ 
 
CASABLANCA, Morocco --May 9, 1990-- Sun Microsystems today announced
that it has appointed Casablanca-based Universal Computer as its first
distributor in Africa.  The agreement will allow Universal Computer to
distribute Sun's hardware and software products throughout Morocco.
Sun France, a subsidiary of the company, will oversee the relationship
with Universal Computer as well as all additional market development
for Sun in Africa. According to Sylvain Idy, distribution director for
Sun France,  "Universal Computer has the expertise to help us gain a
foothold in the  Moroccan market and will be a valuable partner as we
seek to establish  Sun's presence in Africa."  Universal Computer is a
leading distributor of high-technology products in Morocco.  The
company distributes computer hardware and application software,
primarily for CAD, GIS, architecture and commercial markets. According
to Mr. Moussaria, general manager of Universal Computer, "Sun's
arrival in Morocco represents a tremendous opportunity for computer
users in research, industry and management to access the best computer
technology." Idy estimates the size of the computer market in Africa
at approximately 500 million dollars for 1990.  As one of the more
industialized African countries, Morocco's demand for computer
products has increased roughly 25 percent annually, representing a
substantial opportunity for both Sun and Universal Computer. Sun
Microsystems, Inc., headquartered in Mountain View, Calif., is a
leading worldwide supplier of network-based distributed computing
systems, including professional workstations, servers and UNIX
operating system and productivity software.
 
For reader inquiries, telephone 1-800-821-4643 outside California.
Inside California, call 1-800-821-4642.  
Press Contact: Karen Rohack (415) 336-7841

\section{Journals}
The following journals might be of interest to members of the group.
Samples for perusal are available in the AI for Development `library'
(aka filing cabinet) in room F12 of the AI department at 80 South Bridge.
\begin{enumerate}
\item AI Applications in Resource Management  (Darwin Library, K.B.)
\item AI and Society 
\item African Technology Forum
\item Appropriate Technology
\item Computers in Africa
\item Information Technology for Development
\end{enumerate}

\section{Coming Events}
Again, these are all very soon. If you have any information on events
that would be of interest to the group please tell us.
\begin{description}
\item[3rd International Symposium on Artificial Intelligence:]
Applications of Engineering design, Manufacturing and Management in
Industrialised and Developing Countries,
22-26 October 1990, Instituto Tecnologico y de Estudios Superiores de
Monterrey, Monterrey, Mexico.\\
{\bf Contact:} Hugo Terashima, Program Chair, Centro de Intelligencia
Artificial, ITESM, Sucursal do Correos ``J'', C.P. 64849, Monterrey,
N.L. Mexico.\\
e-mail ISAI@TECMTYVM.MTY.ITESM.MX\\
KK also has a copy of the call for papers.
\item[Southern Africa Computer Show:]
 1-3 November 1990, Harare Sheraton, Harare, Zimbabwe.\\
{\bf Contact:} Computers in Africa, 21 Mill Lane, London NW6 1NT,
phone 071 794 5308, FAX 071 435 3607. 
\item[Information Technology in Support of Economic Development:]
9-12 December 1990, Khartoum, Sudan.\\
More details available on this one from ourselves or Computers in
Africa (as above).\\
{\bf Contact:} Professor W. W. Cotterman, Computer Information
Systems, Georgia State University, PO Box 4015, Atlanta, Georgia
30302-4015, USA. 
\end{description}

\section{Contact Personnel}
\begin{tabular}{l|l}
{\bf Software library:} & {\bf Newsletter, overall co-ordination,
meetings:}\\ 
Howard Beck &  Kathleen King\\
Artificial Intelligence Applications Institute &  Department of
Artificial Intelligence \\
University of Edinburgh & University of Edinburgh\\
80 South Bridge & 80 South Bridge\\
Edinburgh EH1 1HN & Edinburgh EH1 1HN\\
031-225-7774 x208 & 031 225 7774 x236\\
hab@uk.ac.ed.aiai & kk@uk.ac.ed.aipna\\
\hline
{\bf Addresses and contacts, funding:} & {\bf Distribution of
information, newsgroup etc:}\\ 
Robert Muetzelfeldt & Olly Morgan\\
Department of Forestry and Natural Resources &  School of Agriculture\\
University of Edinburgh & University of Edinburgh\\
Kings Buildings & Kings Buildings\\
 Mayfield Road &  West Mains Road\\
Edinburgh   EH9 3JU & Edinburgh   EH9 3JU\\
 031 667 1081 & 031 667 1081\\
 R.Muetzelfeldt@uk.ac.edinburgh &  O.Morgan@ed\\
\hline
{\bf Literature resource and bibliography:} &\\
 Ehud Reiter &\\
 Department of Artificial Intelligence &\\
 University of Edinburgh &\\
 80 South Bridge &\\
 Edinburgh EH1 1HN &\\
 031 225 7774 x294 &\\
reiter@uk.ac.ed.aipna &


\end{tabular}

\end{document}



