From: ppc@iam.unibe.ch (PPC '94)
Subject: CFP: Proof Theory, Provability Logic, and Computation (latex)


\documentstyle[12pt]{article}
\pagestyle{empty}
\oddsidemargin  -4.75mm
\evensidemargin -4.75mm
\marginparwidth 2cm
\marginparsep 0pt
\topmargin 0mm
\headheight 0pt
\headsep 0pt
\topskip 0pt        
\textheight 242mm
\textwidth 170mm
\parindent=0pt

\begin{document}
\begin{center}
{\Huge {\boldmath $P^2 C$}{\it\bf '94}}\\[3ex]
{\LARGE\bf International Conference on Proof Theory,}\\[0.7ex]
{\LARGE\bf Provability Logic, and Computation}\\[3.0ex]
{\large March 20--24, 1994}\\
{\large University of Berne, Switzerland}\\[3ex]
{\Large CALL FOR PAPERS}
\end{center}

\bigskip
The central aim of this meeting is to provide a platform for the 
presentation of recent results in the areas of Proof Theory, 
Provability Logic, and Computation where these are interrelated.
More specifically, the main emphasis will be on the following: 
\begin{enumerate}
\item Applications of Proof Theory to Theoretical Computer Science.
\vspace*{-3mm}
\item New developments in Provability Logic related to Computer Science.
\vspace*{-3mm}
\item New challenges from Computer Science for Proof Theory and 
Provability Logic.
\end{enumerate}

The conference is intended for logicians and computer scientists 
interested in the interaction of proof theory
and theoretical computer science.  
The scientific program will consist of invited lectures and
short contributions, which will be selected
from the submitted papers.
All contributions will be refereed for a special issue of the journal 
{\em Annals of Pure and Applied Logic} devoted to the conference, 
where selected papers will be published.

\medskip
An extended abstract (2 pages) of papers to be submitted
should be sent to the address below
not later than January 5, 1994.
The authors will be notified of acceptance for presentation
to the workshop by February 5.
The  complete  paper to be submitted for publication 
should not exceed 15 pages and must be available before April 30, 1994.
The conference fee is SFr 200.-- .

\bigskip 
\begin{tabular}{@{}l@{\hskip 1em}l@{\hskip 3em}l@{\hskip 1em}l}
\multicolumn{2}{@{}l}{\bf Invited Speakers:} \\[0.7ex]
L.\ Beklemishev & (Steklov Institute, Moscow) & 
D.\ de Jongh & (University of Amsterdam) \\
W.\ Buchholz & (University of Munich) & 
F.\ Montagna & (University of Siena) \\
S.\ Buss & (U.C.S.D., San Diego CA) & 
R.F.\ St\"ark & (CIS Munich) \\
R.\ Constable & (Cornell University) & 
A.S.\ Troelstra & (University of Amsterdam) \\
P.\ H\'ajek & (University of Prague)
\end{tabular}

\bigskip 
\begin{tabular}{@{}l@{\hskip 1em}l@{\hskip 3em}l@{\hskip 1em}l}
\multicolumn{2}{@{}l}{\bf Program Committee:} \\[0.7ex]
S.\ Art\"emov & (Steklov Institute, Moscow) \hspace*{5.8mm} & 
S.\ Feferman & (Stanford University) \\
G.\ Boolos & (M.I.T., Cambridge MA) & 
G.\ J\"ager & (University of Berne) \\
E.\ Engeler & (ETH Z\"urich) & 
A.\ Visser & (University of Utrecht)
\end{tabular}

\bigskip
\begin{tabular}{@{}lrl}
Correspondence should be sent to:\\[0.7ex]
PPC '94 &\\
Institut f\"ur Informatik und \hspace*{22.8mm}\\
angewandte Mathematik\\
Universit\"at Bern\\
L\"anggassstrasse 51 & Fax: & +41 31 65 39 65\\
CH--3012 Bern & E-mail: & ppc@iam.unibe.ch
\end{tabular}
\end{document}


