From crabapple.srv.cs.cmu.edu!fs7.ece.cmu.edu!europa.eng.gtefsd.com!howland.reston.ans.net!noc.near.net!news.Brown.EDU!brunix!cs.brown.edu!pvh Mon Aug 16 13:02:04 EDT 1993
Article: 8325 of comp.lang.prolog
Xref: crabapple.srv.cs.cmu.edu comp.lang.prolog:8325
Newsgroups: comp.lang.prolog
Path: crabapple.srv.cs.cmu.edu!fs7.ece.cmu.edu!europa.eng.gtefsd.com!howland.reston.ans.net!noc.near.net!news.Brown.EDU!brunix!cs.brown.edu!pvh
From: pvh@cs.brown.edu (Pascal Van Hentenryck)
Subject: CFP: ICLP-94 (Tex Version)
Message-ID: <1993Aug15.154631.20641@cs.brown.edu>
Keywords: logic programming
Sender: news@cs.brown.edu
Organization: Brown Computer Science Dept.
Date: Sun, 15 Aug 1993 15:46:31 GMT
Lines: 161

\documentstyle{article}
\setlength{\oddsidemargin}{-0.50in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{7.5in}
\setlength{\topmargin}{-0.2in}
\setlength{\headsep}{0.0in}
\setlength{\textheight}{24cm}
\pagestyle{empty}
\hbadness=10000

\begin{document}
\begin{center}
{\Large Call For Papers} \\
$ \ $ \\
{\Large\bf International Conference on Logic Programming (ICLP'94)} \\
$ \ $ \\
S. Margherita Ligure, Italy, 13-18 June 1994 \\
$ \ $ \\
Sponsored by the {\bf Association of Logic Programming}
\end{center}

\parbox{2.3in}{
{\center\bf Program Committee} \\ 
\noindent\begin{tabular}{ll}
Khayri Ali         & Sweden  \\
Maurice Bruynooghe & Belgium \\
Philippe Codognet  & France \\
Yves Deville       & Belgium \\
Herv\'{e} Gallaire & France \\
Chris Hogger       & UK \\
Joxan Jaffar       & USA \\
Giorgio Levi       & Italy \\
Jan Maluszinski    & Sweden \\
Kim Marriott       & Australia \\
Maurizio Martelli  & Italy \\
Lee Naish          & Australia \\
Frank Pfenning     & USA \\
David Poole        & Canada \\
Antonio Porto      & Portugal \\
Raghu Ramakrishnan & USA \\
M. Rodriguez-Artalejo & Spain \\
Gert Smolka	   & Germany \\
V.S. Subrahmanian  & USA \\
Peter Szeredi      & Hungary \\
Evan Tick          & USA \\
Kazunori Ueda      & Japan \\
Pascal Van Hentenryck & USA \\
Peter Van Roy      & France \\
Andrei Voronkov    & Sweden \\
Mark Wallace       & Germany \\
Rong Yang          & UK \\
\end{tabular}
\\
{\center\bf General Chair} \\ 
\begin{tabular}{l}
Maurizio Martelli (Genova) \\
\end{tabular}
%
{\center\bf Program Chair} \\ 
\begin{tabular}{l}
Pascal Van Hentenryck (Brown) \\
\end{tabular}
%
{\center\bf Poster Chair} \\
\begin{tabular}{l}
Lee Naish (Melbourne) \\
\end{tabular}
%
{\center\bf Workshop Chair} \\
\begin{tabular}{l}
Catuscia Palamidessi (Genova) \\
\end{tabular}
%
{\center\bf Publicity Chair} \\ 
\begin{tabular}{l}
Fosca Giannotti (Pisa) \\
\end{tabular}
%
{\center\bf Organizing Committee} \\ 
\begin{tabular}{l}
Rosa Maria Bottino (IMA-CNR) \\
Giorgio Delzanno (DISI) \\
Giuseppe Marino (DIST) \\
Alessandro Messora (DISI) \\
\end{tabular}
%
{\center\bf Local Organization} \\ 
\begin{tabular}{l}
Piera Ponta (CGR) \\
\end{tabular}
}
\ $\; \;$ \
\parbox{4.8in}{
\setlength{\parindent}{0.5cm}

Logic programming originates from the discovery that a subset of
predicate logic could be given a procedural interpretation which was
first embodied in the programming language Prolog. The unique features
of logic programming make it appealing for numerous applications in
artificial intelligence, computer-aided design and verification,
databases, and operations research as well as to explore parallel and
concurrent computing. The last two decades have witnessed substantial
developments in this field from its foundation to implementation,
applications, and the exploration of new language designs.

ICLP'94 is the eleventh international conference on logic programming
and is one of the two major annual international conferences reporting
recent research results in logic programming. The technical program
for the conference will include tutorials, invited lectures, and
presentations of refereed papers and posters. Papers are welcome on
all aspects of logic programming, including, but not limited to:

\begin{quote}
\begin{tabular}{ll}
Applications			& Language design      \\   
Architecture			& Natural language     \\   
Artificial Intelligence 	& Parallelism  \\
Concurrency     		& Programming methodology \\
Constraints			& Proof theory            \\
Databases			& Semantics and foundations \\
Environments			& Static analysis	\\
Higher-order programming	& Theorem Proving       \\
Implementation         		& Types                 \\
\end{tabular}
\end{quote}

Papers must be written in English, must not exceed 15 pages (including
references and figures), and must contain a cover page including the
following: a 200 word abstract, keywords, and postal and electronic
mailing addresses as well as phone numbers and fax numbers of the
responsible author. Submitted papers should not have been previously
published or submitted to any journals or refereed conferences or
workshops. Accepted papers must be presented at the conference. 
Send {\bf six (6) copies} of your submission by {\tt November 15, 1993} to
\begin{quote}
Pascal Van Hentenryck \\
Brown University, Box 1910 \\
Providence, RI 02912 (USA) \\
Email: pvh@cs.brown.edu \\
Phone: +1 401 863 76 34 $\;\;$ Fax: +1 401 863 76 57 
\end{quote}
Authors will be notified of the acceptance or rejection of their
papers by {\tt February 21, 1994}. Final versions of the accepted
papers must be received in camera-ready form by {\tt March 15, 1994.}
The proceedings will be published by {\tt MIT Press.}

ICLP'94 will take place in Santa Margherita Ligure, a small town in
the Italian Riviera close to Genova, the largest city of Liguria.
Close to the conference site is the worldwide famous village of
Portofino, pearl of the Mediterranean Sea and marine natural park; not
far from Santa Margherita is also the pleasant resort area of "Cinque
Terre", consisting of five pictoresque villages on the rocky coast
which can be reached by train or boat only. 
}
\end{document}








